### 2020-11-23 README.Rmd
Initial commit

---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Single cell resolution landscape of equine peripheral blood mononuclear cells reveals diverse cell types including T-bet+ B cells."
```

# Single cell resolution landscape of equine peripheral blood mononuclear cells reveals diverse cell types including T-bet+ B cells.

This repository contains the data analysis scripts and code for our paper:

> Roosheel Patel*, Joy E. Tomlinson*, Thomas J. Divers, Gerlinde R. Van de Walle, Brad Rosenberg, (2020). *Single cell resolution landscape of equine peripheral blood mononuclear cells reveals diverse cell types including T-bet+ B cells*.

Our pre-print is online here:

> Roosheel Patel*, Joy E. Tomlinson*, Thomas J. Divers, Gerlinde R. Van de Walle, Brad Rosenberg, (2020). *Single cell resolution landscape of equine peripheral blood mononuclear cells reveals diverse cell types including T-bet+ B cells* biorxiv, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/10.1101/2020.05.05.077362>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): R Markdown document which includes code to reproduce the figures generated by in our analysis.
  - [:file\_folder: data](/analysis/data): Supplementary data files and metadata used in this analysis

## Additional files required for analysis

For surface annotation of differentially expressed genes, surface protein annotations were pulled from the Cell Surface Protein Atlas website {https://wlab.ethz.ch/cspa/#abstract}.

> Bausch-Fluck D, Hofmann A, Bock T, Frei AP, Cerciello F, et al. (2015) *A Mass Spectrometric-Derived Cell Surface Protein Atlas*. PLoS One <https://doi.org/10.1371/journal.pone.0121314>

### Licenses

**Code :** MIT License, Copyright (c) 2020 Roosheel Patel

The organization of this repository was adapted from the rrtools research compendium structure

> Marwick, B. (2017). Computational reproducibility in archaeological research: Basic principles and a case study of their implementation.*Journal of Archaeological Method and Theory*, 24(2), 424-450.<https://doi.org/10.1007/s10816-015-9272-9>